---
title: "Sesion 14: Reduccion de dimensionalidad"
author: "Juan Carlos Martinez-Ovando"
date: "Primavera 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### **Resumen y objetivos**

* Revisaremos los fundamentos del analisis de factores y el papel que juegan las variables latentes en su formulacion.

* dada la inclusion de variables latentes, revisaremos como se implementa el _aprendizaje bayesiano_ en esta clase de modelos.

* Presentaremos alguans ilustraciones.

## B. Analisis de factores

### B-1) Datos

```{r}
rm(list=ls())
setwd("/home/jcmo/JCMO.Trabajo/@Mis.Cursos/2017-A_Aprendizaje Estadistico/sesiones")

data <- read.csv("FA/MCMC_FactAnal.csv",header=TRUE, sep=",",na.strings="NA", dec=".", strip.white=TRUE)
summary(data)
head(data)
```

### B-2) Aprendizaje frecuentista

```{r}
fa.1 <- factanal(~., factors = 4, rotation = "varimax", data = data)
fa.1

print(fa.1, digits = 3, cutoff = .000001)
```

### B-3) Aprendizaje bayesiano

Emplea MCMC.

```{r}
require("MCMCpack")
help(MCMCfactanal)
fa.2 <- MCMCfactanal(~x1+x2+x3+x4+x5+x6+x7+x8, 
        factors = 2, data = data, burnin = 1000, 
        mcmc = 10000, thin = 10, verbose = 0, seed = NA,
        lambda.start = NA, psi.start = NA, l0 = 0, L0 = 0, 
        a0 = 0.001, b0 = 0.001, store.scores = FALSE, std.var = TRUE)
summary(fa.2)
plot(fa.2)
heidel.diag(fa.2)
```

```{r}
# Extracting elements of the output. 

### First, the loadings (Lambda). 

loadings <- summary(fa.2)$statistics[1:16]
loadings

# Factor 1 item loadings (items x1, x2, x3, x4).

factor1.item.loadings <- loadings[c(1,3,5,7)]
factor1.item.loadings

# Factor 2 item loadings (items x5, x6, x7, x8).

factor2.item.loadings <- loadings[c(10,12,14,16)]
factor2.item.loadings

# Extract the 95% Credible Interval estimates of each loading.

ci.loadings <- summary(fa.2)$quantiles[1:16,c(1,5)]
ci.loadings

# Factor 1 item loading intervals (items x1, x2, x3, x4).

factor1.ci.loadings <- ci.loadings[c(1,3,5,7),]
factor1.ci.loadings

# Factor 2 item loading intervals (items x5, x6, x7, x8).

factor2.ci.loadings <- ci.loadings[c(10,12,14,16),]
factor2.ci.loadings

### Extract the uniquenesses (Psi). 

uniquenesses <- data.frame(summary(fa.2)$statistics[17:24],
                           names(data))
names(uniquenesses)[1] <- "uniquenesses"
uniquenesses

# Extract the 95% Credible Interval estimates of each uniqueness.

ci.unique <- summary(fa.2)$quantiles[17:24,c(1,5)]
ci.unique

# Calculate the Communalities for each item.

communalities <- data.frame(1 - uniquenesses[,1], names(data))
names(communalities)[1] <- "communalities"
communalities
```

```{r}
################################################################################

# Here constraining the loadings (Lambda) so that the first four items load exclusively on 
# factor 1 and not on factor 2; as well as ensuring the second four items load on factor 2 
# and not on factor 1; for instance, item "x1" is constrained on factor 2 to have a loading 
# of zero [e.g. x1=c(2,0)]. One could say this is a more 'confirmatory' strategy; which seems 
# appropriate when taking a Bayesian approach (i.e. the specification of a prior, which 
# is necessary; indicates an exploratory factor analysis was done previoiusly or previous
# research has established the structure of the measure). 

fa.3 <- MCMCfactanal(~x1+x2+x3+x4+x5+x6+x7+x8, 
        factors = 2, data = data, 
        lambda.constraints=list(x1=list(2,0), 
        x2=c(2,0), x3=list(2,0), x4=c(2,0),
        x5=list(1,0), x6=c(1,0), x7=list(1,0), x8=c(1,0)),
        burnin = 1000, mcmc = 10000, thin = 10, verbose = 0,
        seed = NA, lambda.start = NA, psi.start = NA, 
        l0 = 0, L0 = 0, a0 = 0.001, b0 = 0.001, 
        store.scores = FALSE, std.var = TRUE)
summary(fa.3)

# Notice in the summary (above) the loadings are substantially greater than the previous MCMCfactanal,
# and the items are not allowed to cross-load (i.e. load on a factor they should not). In the situation 
# where items are thought to load on multiple factors, constraints can be used to specify how 
# each item loads on each factor (positive / negative; or magnitude of the loading for each factor). 

# Below, running the Heidel diagnostic test (from package 'coda') confirms stationarity was achieved. 

heidel.diag(fa.3)

# Also, the densities for the loadings are more normally distributed. 

plot(fa.3)


ls()

# NOTE: The 'MCMCpack' package also contains a function for doing MCMC factor analysis 
# with mixed data (i.e. continuous and ordinal variables).

help(MCMCmixfactanal)


################################################################################

# Martin, A. D., Quinn, K. M., & Park, J. H. (2010). Package 'MCMCpack'. 
# Package Reference Manual available at:
# http://cran.r-project.org/web/packages/MCMCpack/MCMCpack.pdf

# Plumer, M., Best, N., Cowles, K, & Vines, K. (2010). Package 'coda'. 
# Package Reference Manual available at:
# http://cran.r-project.org/web/packages/coda/coda.pdf





# End: 7 Jan. 2010.

```

## Referencias

* **Alpaydin** -- Capitulo 6 

* **Barber** -- Capitulo 15 

* **Hastie et al** -- Seccion 14.5


